---
title: "GGPlot Exercises"
author: "Xuhao Dong"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(scales)
library(RColorBrewer)
```

## Introduction

This document takes us through some common charting challenges with `ggplot`. Much of this
code is reinforced or introduced in the "Data Visualization" chapter of _Algorithms for Data Science_.

When you submit your work, please submit both your RMD file and the knitted HTML file. Save the 
data for this assignment in the same folder as the RMD and the reads should work. 

```{r Read in data}

gd <- read_tsv("grocery_data.txt")
month.summary <- read_tsv("month_summary.txt")
dept.summary <- read_tsv("dept_summary.txt")
mosaic.data <- read_tsv("counts_for_mosaic.txt")

```

## Problem 1 - Sales by month by department

Things to add to this chart: 

* Divide sales by 1000
* Add some transparency to the points with `alpha`
* Add axis labels using `labs`
* Center the title, rather than it being left-aligned

```{r prob-1}

gd <- gd %>%
    mutate(month = as.factor(month),
           departmentName = reorder(departmentName,ownerSales,mean))
head(gd,10)

# Here's a basic one:
plot <- ggplot(gd, aes(x=(sales = ownerSales / 1000),y=departmentName)) + 
  geom_point(alpha = 1/10) 

plot <- plot + labs(x = "Owner Sales", y = "Department Name")

plot <- plot +
  ggtitle("Sales by month by department") +
  theme(plot.title = element_text(hjust = 0.5))

print(plot)
```

## Problem 2 - add median lines

Add these lines to the plot using geom_segment.

* That function/layer takes four values
* x, xend, y, and y.end. You can get y.values with
* something like `as.numeric(departmentName)-0.5` (or plus)
* The x's should be the median values


```{r median-lines}

medians <- gd %>% 
  group_by(departmentName) %>%
  summarise(med.val = median(ownerSales/1000))

medians

ggplot(gd, aes(x=ownerSales/1000,y=departmentName)) + 
  geom_point() +
  geom_segment(aes(x = med.val, y = as.numeric(departmentName)-0.5, xend = med.val, yend = as.numeric(departmentName)+0.5, colour = "segment"), data = medians)

```

## Problem 3 - line chart plotting the density

Things to add:

* a theme (could use `theme_minimal`)
* label each line using `geom_text`
* Each label should read "Bandwidth = x"
* Make sure labels don't overlap the line (use `check_overlap=T` within `geom_text` to accomplish this)
* Each label should match the color of the line its labeling

```{r prob-3}

p3.size <- 4
p3.cols <- c("gray15","gray40","gray75")
p <- ggplot(gd,
            aes(x=ownerSales/1000)) + 
  ylab("Density") + 
  geom_line(stat="density",col=p3.cols[1],size=1.25,bw=1) + 
  geom_line(stat="density",col=p3.cols[2],size=1.25,bw=5) + 
  geom_line(stat="density",col=p3.cols[3],size=1.25,bw=25) + 
  scale_x_continuous(label=dollar) + 
  xlab("Monthly Sales (000)") + 
  geom_text(aes(x=20,y=0.05),
            label= 'Bandwidth = 1', 
            size=p3.size, check_overlap = T, col=p3.cols[1]) + 
  geom_text(aes(x=20.5,y=0.03),
            label= 'Bandwidth = 5',
            size=p3.size, check_overlap = T, col=p3.cols[2]) + 
  geom_text(aes(x=75,y=0.009), 
            label='Bandwidth = 25',
            size=p3.size, check_overlap = T, col=p3.cols[3]) +
  theme_minimal()

print(p)


```

## Problem 4 - use `gd` to create a density plot of sales by department. 

Things to add:

* First, make a density plot of sales (using `geom_density`). 
* Then facet that plot by department using `facet_wrap`
* Change the x axis tick marks so they are easily readable

```{r prob-4}

den <- ggplot(gd, aes(ownerSales/1000, ..count..)) +
  geom_density(adjust = 5) + 
  ylab("Density") +
  scale_x_continuous(label=dollar) + 
  xlab("Monthly Sales (000)")

den2 <- den + facet_wrap(vars(departmentName))

den3 <- den2 + theme(axis.text.x = element_text(face="bold", color="#993333", 
                           size=7, angle=45),
          axis.text.y = element_text(face="bold", color="#993333", 
                           size=7, angle=45))
den3
```

## Problem 5 - violin plots! Use gd, build violin plots of ownerSales by department. 

Things to add:

* add quantile lines, use nice labels,
* re-orient chart with departments on the y-axis 
* add a vertical line for the overall mean and change its color
* add an annotation to the mean line with the mean sales


```{r prob-5}

mean <- gd %>% 
  pull(ownerSales) %>% 
  mean() %>%
  signif(6)

v <- ggplot(gd, aes(ownerSales,departmentName)) +
  xlab("Owner Sales")+
  ylab("Department Name")+
  geom_violin(scale='width',trim=FALSE) + 
  geom_boxplot(width=0.1) +
  geom_vline(xintercept=mean,size=1.5, color="red")+
  geom_text(aes(x=mean, label=paste0("Mean\n",mean), y=1.9))

v

```


## Problem 6 - Mosaic charts

Things to add:

* Using mosaic.table, create a mosaic plot
* Color the mosaicplot using the default gray pallette. 


```{r prob-6}

mosaic.table <- mosaic.data %>% 
  spread(Segment,count) %>% 
  rename(Dept=DepartmentName) %>% 
  arrange(Primary) %>% 
  as.data.frame
  
head(mosaic.table,10)
# Let's do some renaming
row.names(mosaic.table) <- mosaic.table$Dept
mosaic.table$Dept <- NULL

rownames(mosaic.table)[rownames(mosaic.table)=="SUPPLEMENTS"] <- "SUPP."
rownames(mosaic.table)[rownames(mosaic.table)=="REF GROCERY"] <- "REF GROC"
rownames(mosaic.table)[rownames(mosaic.table)=="PACKAGED GROCERY"] <- "PKG GROC"

mosaicplot(mosaic.table, color = TRUE, main = "Department Shopping Frequency")
```

