---
title: "Survey Clustering"
author: "Xuhao Dong"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r includes, echo=F, warning=F, message=F}
library(cluster)
library(tidyverse)
library(pvclust)
library(knitr)
```

## Clustering

This assignment asks you to cluster the Washington survey data explored in the 
$\chi^2$ and ANOVA assignments. A full description of that data is in those 
markdown files. I've included the appendix data description at the bottom. Since
those assignments require EDA, you _don't_ have to do an EDA in this assignment.

```{r data-input}
d <- readr::read_tsv("washington_survey_data.txt")
```

## Demo Clustering

In this section, perform a cluster analysis using the non-value fields: 

* age
* gender
* engagement
* account.age

Choose a number of clusters and use `pam` to cluster the data using these fields. Briefly 
describe the clusters. How evenly do your clusters align with the regions?  


```{r demo clustering}
#names(d) <- c("ID", "Age", "Gender", "Engagement",
              # "Mem.Edu", "Zip", "Channel", "Progressivism",
              # "Harm", "Fair", "In.Group", "Authority",
              # "Purity","Account_Age", "Region","Public_Sector",
              # "Sustainability","Localism", "Pub_Priv", "Experience",
              # "Underpaid", "Focal_Value","Local_Support")


data.for.clust <- d %>% select(age,gender,engagement,account.age)

clust <- 
  pam(x=data.for.clust,
    k=3)
#summary(clust)
#table(clust$clustering)
d<- d %>% mutate(member= clust$clustering)

a <- d %>% count(member,region)
kable(a)

```

Most people live in western Washington Metro Area in each cluster. The next most popular area to live in is Thurston. This seems to be constant across all clusters. Cluster 2 has the greatest population of 950 and Cluster 3 has a population of 674. 

## Values Clustering

Similarly to the previous section, perform a cluster analysis, this time on the
values questions: 

* fair
* harm
* in.group
* authority
* purity
* public.sector
* sustainability
* localism

After you've built your clusters, report the following information on each 
cluster: 

1. Predominant region
1. Average age and account age
1. Most common focal value
1. Mean results on the questions of `pub.greater.priv`, `experience.more.important`,
and `teachers.underpaid`

```{r value clustering}
data.for.clust1 <- d %>% select(fair,harm,in.group,authority,purity,public.sector,sustainability,localism)

clust <- 
  pam(x=data.for.clust1,
    k=5)
#summary(clust)
#table(clust$clustering)
d1 <- d %>% mutate(membership = clust$clustering)

x <- d1 %>%  group_by(membership) %>% summarize("Age" =mean(age), "Account_Age" = mean(account.age), "Priv" =mean(pub.greater.priv), "Experience" =mean(experience.more.important), Underpaid = mean(teachers.underpaid))
kable(x)

y<- d1 %>% count(membership,region)
kable(y)
   
z<- d1 %>% count(membership,main.focal.value)
kable(z)
                                     
```
The analysis cluster data into 5 group. In cluster 1, a mean customer age of 43.65, an average account age of 12.24 years, the most predominant region is the Western Washington Metro area. This clusters most common value is Education. Also a mean of 3.08 on greater public priv scoired, a mean score of 4.0 on experience and a mean score of 4.53 on if teachers are underpaid. 

Cluster 2 has a mean age of 49.26, mean account age of 14.22 years, and are predominately located in the W WA Metro. The most common focal value is environment. The clusters mean score for greater public priv was 3.37, mean score for experience was 3.87 and mean score for underpaid teachers was 5.31. 

Cluster 3 has a mean age of 53, mean account age of 14.91 years, and are also most located in the W WA Metro with Thurston being a close second (254 and 215 respectively). They are mostly interested in Education, Health, Hunger and Poverty. They rank third in greater public priv with a mean score of 3.24, 4th in experience with a mean score of 3.97, and second in underpaid teachers with a mean score of 5.04.

Cluster 4 is the oldest age group, with an average age of 55.07, they have held their accounts on average 13.1 years, and primarily live in the W WA Metro and Thurston area (152 and 140). The primary focal interest is health with 78, and education with 76. They have a mean greater public priv score of 3.29, a mean score for experience of 4.14, and a mean underpaid teacher score of 4.91.

Cluster 5 has mean age of 52.04, mean account age of 14.12, and primarily live in W WA Metro and Thurston (65 and 61). They cares about health and law enforcement with 50 and 45 votes each. They have the lowest score of greater public priv with an average score of 2.58. This cluster values experience the most with and average score of 4.18 and they scored on average the lowest for underpaid teachers with mean score of 4.05. 

# Appendix: Full Data Description
A financial institution in Washington has become concerned that their current membership 
base is not well-aligned with their corporate values. Through that concern they 
realized that don't actually understand their membership's values very well. They 
surveyed 2,421 members to shed light on the issue. 

The heart of the survey was the Moral Foundations Theory of Jonathan Haidt. Members 
were surveyed on the Moral Foundations Questionnaire, which you should take so you 
understand the test. Survey respondents were scored on the five foundations as well 
as a single-number summary, Progressivism. 

The financial institution values Localism, Sustainability, and Education. These aspects 
of member's values were assessed in the survey as well. Localism and Sustainability used
validated scales and thus can be summarized via a single score, where higher values indicate
greater support for the values. Education is summarized by the following three questions, 
which we do not have evidence can be combined into a single score:

* In general, public schools provide a better education than private schools.
* Public school teachers are underpaid.
* Experience is more important than education in determining success in life.
These questions were evaluated on a 1 to 6 scale where 1 indicated "Strongly Disagree" and 
6 indicated "Strongly Agree". 

Finally, we have information on the member that can be used to understand variation 
in their values. 

The data consists of the following columns:

* ID: a unique identifier for the survey respondent.
* age: the age of the respondent.
* gender: gender was evaluated with robust scale and collapsed into male/female/other for 
  those whose gender identity was not male or female.
* engagement: three categories of engagement with the financial institution.
* mem.edu: the self-reported education level of the member with the following scale:
* zip: the member zip code. 
* channel: how the member joined the financial institution. Options are "Loan" if they joined 
  via an auto loan, "Branch" if they joined at a branch and other for online or unknown. 
* progressivism/harm/fair/in.group/authority/purity: The MFQ results.
* account.age: the age of the member's account, in years. 
* region: The region of Washington the member lives in. May be easier to work with than zip.
* public.sector: has the person ever been a public employee?
* sustainability/localism: Scores on the validated scales. Higher values indicate greater
  support for the value.
* pub.greater.priv/experience.more.important/teachers.underpaid: The responses to the 
  education questions above. 
* main.focal.value: Respondents were asked, "Below is a list of broad areas to which people 
  often dedicate their volunteer or philanthropic efforts. From this list, please select the 
  most important to you. If an area of particular importance is missing, please let us know 
  about it in the space for 'other.'" This column holds the respondents' answer to that question.
* support.of.focal.value: Respondents were given an opportunity to indicate how they 
  supported their focal value. Those responses were collapsed into a single score, where 
  a higher value indicates more support.





