---
title: "Future Predicted Value"
author: "Xuhao Dong"
date: "2021-12-05" # fix the date!
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=F)
```

## Future Predicted Value

Future predicted value is the estimate of how much someone will spend in the future with a 
company. In this report we calculate FPV for the Wedge in 2015 using data from before that 
year. Our report has the following sections: 

1. Variable selection: In this section we describe the variables that we use in the model.
1. Model formulation: This section describes the model itself. 
1. Estimates of model accuracy: We run a quick analysis of model accuracy in the final section.

```{r includes}
# Feel free to adjust to the ones you need.
library(tidyverse)
library(scales)
library(reshape2)
library(DBI)
library(broom)
library(greybox)
```


```{r db-hookup}
# Change to your path
path.to.db <- paste0("/Users/steviodong/Desktop/Applied Data Analytics/TWP/final/")

# Creating the connection to the DB. Similar to 
# a cursor. 
con <- dbConnect(RSQLite::SQLite(),
                 dbname=paste0(path.to.db,"thewedge_part3.db"))

# Listing the tables
dbListTables(con) # uncomment to see table names

# And creating connections to the two tables. 
owner.ym <- tbl(con, "owner_sales_by_yearmonth")

```

### Variable Selection

```{r variable-selection}
# Use this space to create the data set you'll feed into `lm` I've made a simple one for you.
# I've done this in two different ways so you can see a couple of techniques.
# First way uses "pivot_wider"
md <- owner.ym %>% 
  filter(card_no != "3",
         between(year,2013,2015)) %>% 
  group_by(card_no,year) %>%
  summarize(sales=sum(sales,na.rm=T)) %>% 
  ungroup() %>% 
  collect 

md <- md %>% 
  pivot_wider(id_cols=c("card_no"),
              names_from=year,
              names_prefix="sales_",
              values_from=sales)

# Second builds a data set up from the ground
md2 <- owner.ym %>% 
  filter(card_no != "3",
         between(year,2013,2015)) %>% 
  distinct(card_no) %>% 
  collect 

# Now start bringing in our covariates. 
md2 <- left_join(md2,
                owner.ym %>% 
                  filter(year==2013) %>% 
                  group_by(card_no) %>% 
                  summarize(sales_2013 = sum(sales)) %>% 
                  collect,
                by="card_no") 

md2 <- left_join(md2,
                owner.ym %>% 
                  filter(year==2014) %>% 
                  group_by(card_no) %>% 
                  summarize(sales_2014 = sum(sales)) %>% 
                  collect,
                by="card_no") 

md2 <- left_join(md2,
                owner.ym %>% 
                  filter(year==2015) %>% 
                  group_by(card_no) %>% 
                  summarize(sales_2015 = sum(sales)) %>% 
                  collect,
                by="card_no") 


```
```{r split data}
#This takes number of rows and divides by two for a 10/90 split
bound <- floor((nrow(md)*0.1))

#This randomly shuffles the entire dataset
md <- md[sample(nrow(md)),]

#Create a new dataframe called md.train that contains record #1 to "bound"
md.train <- md[1:bound, ]

#Create a new dataframe called md.test that contains record "bound" to the end of the dataframe
md.test <- md[(bound+1):nrow(md), ]

```



I first created a bound and take the number of row and split into 10/90, then shuffle the dataset and put them into two new dataframe md.train and md.test.


### Model Formulation

```{r modeling}
# A simple example model
lm.1 <- lm(sales_2015 ~ sales_2013 + sales_2014,
           data= md.train,
           subset=sales_2015 <= 15000)
plot(lm.1)
# Of course, you'd never just dump a summary table into a document!
print(tidy(lm.1))
summary(lm.1)
```
Q2:Do a short write up of your model. What terms are significant? What's the R^2? What's the 
residual standard error and how would you interpret it?

The sales of 2014 is very significant to the model, an adjusted r-squared of 77.60% reveals that 77.60% of the data fit the regression model. Residual Standard Error is measure of the quality of a linear regression fit. In our example, the actual sales in 2015 can deviate from the true regression line by approximately 1491 dollars, on average. 
 


### Model Accuracy

```{r model-accuracy}
#Place the prediction variable back in the test dataset
md.test$pred <- predict(lm.1,newdata = md.test)

MAPE(md.test$sales_2015,md.test$pred)

mean(abs((md.test$sales_2015-md.test$pred)/md.test$sales_2015),na.rm=TRUE)*100

```

<!--

As the readme indicates, do a quick check of model accuracy on a holdout sample that
isn't use for fitting.

-->

